Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-04-23 main.Rmd
proof reading

Consider a linear granularity $G'$ such that $G' = Alter^{2}_{(1,-2)}(B, Group_7(B))$, which implies it is made up of shrinking every 1st granule of $Group_7(B)$ by $2$ granules.<!-- $$G'= Alter^{2}_{(1,-1)}(BG, Group_3(B))$$ and --><!--This implies that $G'$ is made up by shrinking every 1st granule of $Group_3(B)$ by $1$ granule and $H'$ is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules. Number of granules of $G'$ and $H'$ in each period of $B$ is 2 but the number of granules of $B$ in each of those granules are different.--> Then $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$. Here, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$. Hence using \ref{def:quasicircular}, then we will have:

@@

<!-- **Discussion**:  -->
If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$ (from definition \ref{def:periodical})<!-- - $B \trianglelefteq G'$ --> <!-- and,  number of granules of $B$ in each granule of $G'$ is not a constant (irregular mapping) -->. Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$. The idea here is if we know the composition of each of the granules of $G'$ in terms of granules of $B$ for one period, we can find the composition of any granule of $G'$ beyond a period since the "pattern" repeats itself along the time domain due to the periodic property. The periodic property also ensures that $\vert T_w \vert = \vert T_{w\mod R'} \vert$ since every $w^{\text{th}}$ and $(w+R')^{\text{th}}$ granule of $G'$ will have the same number of granules of $B$. The term $\sum_{w=0}^{k-1}\vert T_{w}\vert$ in Definition \autoref{def:quasicircular} denotes the number of granules of $B$ till the $(k-1)^{\text{th}}$ granule of $G'$. Since $\vert T_w \vert = \vert T_{w\mod R'} \vert$, the number of granules of $B$ till the $(k-1)^{\text{th}}$ granule of $G'$ becomes $\sum_{w=0}^{k-1}\vert T_{w \mod R'}\vert$ in Definition \autoref{def:quasicircular}.

@@

<!--Periodic and Quasi-periodic behavior can be defined by a repeating pattern except for a finite number of granules that can be seen as discontinuity points in the granularity in case of quasi-periodic behavior.--> Aperiodic time granularities are the ones which can not be specified as a periodic repetition of a pattern of granules. Most public holidays repeat every year, but there is no finite (or reasonably small) period within which their behavior remains constant. A classic example can be that of Easter, whose dates repeat only after 5,700,000 years.<!--U.S labor day is the first Monday in September and U.S. Memorial Day is the last day in May. --> In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend. Examples of aperiodic granularity may also include school holidays or a scheduling event.<!--that might cover the first and third Monday of the month between June and October, except for state holidays.--> All of these are recurring events, but with non-periodic patterns. As such, plausible $P_i$ from \autoref{tab:quasi} could be possibly infinite for aperiodic granularities.

@@

Consider a school semester which always lasts for 18 weeks and 2 days, starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 9 days. Let the linear granularities $M_1$ and $M_2$ denote in-session semester period and semester break periods respectively. Both $M_1$, $M_2$ and $M = M_{1}\bigcup M_{2}$ denoting the semester week type are aperiodic with respect to days ($B$) or weeks ($G$). Hence $A_{B, M}$ denoting day of the semester week type would be an aperiodic cyclic granularity. This is because the placement of the semester within an year would vary across years (\autoref{fig:aperiodic-example} c). \autoref{fig:aperiodic-example} b shows the stack display of time with granules representing same categories (semester break/in-session) stacked on top of one another. \autoref{fig:aperiodic-example} (a) representing the linear display of time is useful in manifesting how the categories would be determined with respect to the bottom granularities using Definition \ref{def:aperiodic}. It is interesting to note here that $Q_{H, M}$ denoting semester week of the semester week type would be a quasi-circular granularity since the distribution of semester weeks within a semester is assumed to remain constant over years. Here $Q_{H, M}$ with a period of 18.3 weeks will have irregular mapping since each semester week types consists of different number of semester weeks.

@@

\begin{equation} \label{eq:eq7}
\begin{split}
C_{G_l,G_m}(z) & = C_{G_l,G_{l+1}}(z) + P(l, l+1)(C_{G_{l+1},G_m}(z)) \\
  & = C_{(G_l,G_{l+1}}(z) + P(l, l+1)[ C_{G_{l+1},G_{l+2}}(z) + P(l+1, l+2)]( C_{G_{l+2}, G_m}(z)) \\
  & = C_{G_l,G_{l+1}}(z) + P(l, l+1)(C_{G_{l+1},G_{l+2}}(z)) + P(l, l+1)P(l+1, l+2)(C_{G_{l+2}, G_{m}})\\
  & = C_{G_l,G_{l+1}}(z) + P(l, l+1)(C_{G_{l+1},G_{l+2}}(z)) + P(l, l+2)(C_{G_{l+2},G_{l+m}}(z))\\
  &\vdots\\
  & = \sum_{i=0}^{m - l - 1} P(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z))\\
\end{split}
\end{equation}

@@

This case is similar to Equation \ref{eq:multifromsingle-quasi2}. Single-order-up quasi-circular granularity ($C_{G_{l},G_{m'}}$) could be obtained from multiple-order-up quasi-circular granularity ($C_{G_l,G_{m}}$) and single/multiple-order-up circular granularity $C_{m',m}$ using \ref{eq:multifromsingle-quasi2}.

\begin{equation} \label{eq:multifromsingle-quasi3}
\begin{split}
C_{G_{l},G_{m'}}(z) & = C_{G_l,G_{m}}(z) - \sum_{w=0}^{C_{m',m}(z) -1}(\vert T_{w} \vert)\\
\end{split}
\end{equation}

where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{l}$, $\vert T_w \vert$ is the cardinality of set $T_w$.
<!-- as they promote the exploratory analysis through visualization. -->

>>>>

Consider a linear granularity $G'$ such that $G' = Alter^{2}_{(1,-2)}(B, Group_7(B))$, which implies it is made up of shrinking every 1st granule of $Group_7(B)$ by $2$ granules.<!-- $$G'= Alter^{2}_{(1,-1)}(BG, Group_3(B))$$ and --><!--This implies that $G'$ is made up by shrinking every 1st granule of $Group_3(B)$ by $1$ granule and $H'$ is made up of shrinking every 1st granule of $Group_3(B)$ by $2$ granules. Number of granules of $G'$ and $H'$ in each period of $B$ is 2 but the number of granules of $B$ in each of those granules are different.--> Then $Q_{B, G'}$ is a repetitive categorization of time, similar to circular granularities, except that the number of granules of $B$ is not necessarily the same across different granules of $G'$. Here, $T_0 = \{0, 1, 2, 3, 4, 5, 6\}$ and $T_1 = \{7, 8, 9, 10, 11\}$. Hence using Defintion \autoref{def:quasicircular} we will have:

@@

<!-- **Discussion**:  -->
If linear granularity $G'$ is periodical with respect to $B$ with irregular mapping, then there exist $R'$, $P' \in \mathbb{Z}_+$ such that if $G'(w) = \bigcup_{z \in T_w}B(z)$ then $$G'(w) = \bigcup_{z \in T_w \mod R'}B(P'*\lfloor w/R' \rfloor + z)$$ (from Definition \ref{def:periodical})<!-- - $B \trianglelefteq G'$ --> <!-- and,  number of granules of $B$ in each granule of $G'$ is not a constant (irregular mapping) -->. Here $w \mod R'$ represents the index that must be shifted to obtain $G'(w)$. The idea here is if we know the composition of each of the granules of $G'$ in terms of granules of $B$ for one period, we can find the composition of any granule of $G'$ beyond a period since the "pattern" repeats itself along the time domain due to the periodic property. The periodic property also ensures that $\vert T_w \vert = \vert T_{w\mod R'} \vert$ since every $w^{\text{th}}$ and $(w+R')^{\text{th}}$ granule of $G'$ will have the same number of granules of $B$. The term $\sum_{w=0}^{k-1}\vert T_{w}\vert$ in Definition \autoref{def:quasicircular} denotes the number of granules of $B$ till the $(k-1)^{\text{th}}$ granule of $G'$. Since $\vert T_w \vert = \vert T_{w\mod R'} \vert$, the number of granules of $B$ till the $(k-1)^{\text{th}}$ granule of $G'$ becomes $\sum_{w=0}^{k-1}\vert T_{w \mod R'}\vert$ in Definition \autoref{def:quasicircular}.

@@

<!--Periodic and Quasi-periodic behavior can be defined by a repeating pattern except for a finite number of granules that can be seen as discontinuity points in the granularity in case of quasi-periodic behavior.--> Aperiodic time granularities are the ones which can not be specified as a periodic repetition of a pattern of granules. Most public holidays repeat every year, but there is no reasonably small period within which their behavior remains constant. A classic example can be that of Easter, whose dates repeat only after 5,700,000 years.<!--U.S labor day is the first Monday in September and U.S. Memorial Day is the last day in May. --> In Australia, if a standard public holiday falls on a weekend, a substitute public holiday will sometimes be observed on the first non-weekend day (usually Monday) after the weekend. Examples of aperiodic granularity may also include school holidays or a scheduling event.<!--that might cover the first and third Monday of the month between June and October, except for state holidays.--> All of these are recurring events, but with non-periodic patterns. As such, plausible $P_i$ from \autoref{tab:quasi} could be possibly infinite for aperiodic granularities.

@@

Consider a school semester which always lasts for 18 weeks and 2 days, starting with one week of orientation followed by an in-session period of 6 weeks, a semester-break of 1 week and again an in-session period of 7 weeks and 1 week study break before final exams which continue for 9 days. Let the linear granularities $M_1$ and $M_2$ denote in-session semester period and semester break periods respectively. Both $M_1$, $M_2$ and $M = M_{1}\bigcup M_{2}$ denoting the semester week type are aperiodic with respect to days ($B$) or weeks ($G$). Hence $A_{B, M}$ denoting day of the semester week type would be an aperiodic cyclic granularity. This is because the placement of the semester within an year would vary across years. \autoref{fig:aperiodic-example} b shows the stack display of time with granules representing same categories (semester break/in-session) stacked on top of one another. \autoref{fig:aperiodic-example} (a) representing the linear display of time is useful in manifesting how the categories would be determined with respect to the bottom granularities using Definition \ref{def:aperiodic}. It is interesting to note here that $Q_{H, M}$ denoting semester week of the semester week type would be a quasi-circular granularity since the distribution of semester weeks within a semester is assumed to remain constant over years. Here $Q_{H, M}$ with a period of 18.3 weeks will have irregular mapping since each semester week types consists of different number of semester weeks.

@@

\begin{equation} \label{eq:eq7}
\begin{split}
C_{G_l,G_m}(z) & = C_{G_l,G_{l+1}}(z) + P(l, l+1)C_{G_{l+1},G_m}(z)\\
  & = C_{G_l,G_{l+1}}(z) + P(l, l+1)( C_{G_{l+1},G_{l+2}}(z) + P(l+1, l+2)C_{G_{l+2}, G_m}(z)) \\
  & = C_{G_l,G_{l+1}}(z) + P(l, l+1)C_{G_{l+1},G_{l+2}}(z) + P(l, l+1)P(l+1, l+2)C_{G_{l+2}, G_{m}}\\
  & = C_{G_l,G_{l+1}}(z) + P(l, l+1)C_{G_{l+1},G_{l+2}}(z) + P(l, l+2)C_{G_{l+2},G_{l+m}}(z)\\
  &\vdots\\
  & = \sum_{i=0}^{m - l - 1} P(l, l+i)C_{G_{l+i},G_{l+i+1}}(z)\\
\end{split}
\end{equation}

@@

Single-order-up quasi-circular granularity could be obtained from multiple-order-up quasi-circular granularity and single/multiple-order-up circular granularity using Equations \ref{eq:multifromsingle-quasi1} and \ref{eq:multifromsingle-quasi2}.

<!-- \begin{equation} \label{eq:multifromsingle-quasi3} -->
<!-- \begin{split} -->
<!-- C_{G_{l},G_{m'}}(z) & = C_{G_l,G_{m}}(z) - \sum_{w=0}^{C_{m',m}(z) -1}(\vert T_{w} \vert)\\ -->
<!-- \end{split} -->
<!-- \end{equation} -->

<!-- where, $T_w$ is such that $G_{m'}(w) = \bigcup_{z \in T_w}G_{l}$, $\vert T_w \vert$ is the cardinality of set $T_w$. -->
<!-- as they promote the exploratory analysis through visualization. -->
